---
title: "Group 3 Final Project Script"
author: "Authors: Ryan Charter, David Chen, Ria Krishna, Fatim Majumder, Mia Roselinsky"
date: "Date: 2022-12-04"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{css, echo = FALSE}
h1, h3{
  text-align: center;
}
```

# Import Libraries 
```{r, message = FALSE}
library(tidyr)
library(dplyr)
library(stringr)
library(ggplot2)
library(plotly)
library(lubridate)
library(ggmap)

fulldata <- read.csv("fulldata.csv")
```

# Data Cleaning

#### Change the dtype of variables
```{r, warning = FALSE}
fulldata$srch_destination_latitude <- as.numeric(fulldata$srch_destination_latitude)
fulldata$srch_destination_longitude <- as.numeric(fulldata$srch_destination_longitude)
fulldata$hour <- hour(format(fulldata$date_time, format = "%H:%M:%S"))
fulldata$date_time <- as.Date(fulldata$date_time)
fulldata$srch_ci <- as.Date(fulldata$srch_ci)
fulldata$srch_co <- as.Date(fulldata$srch_co)
fulldata$orig_destination_distance <- as.numeric(fulldata$orig_destination_distance)
fulldata$prop_starrating_cat <- as.factor(fulldata$prop_starrating)
```

#### Change binary variables into factors
```{r}
fulldata$prop_is_branded[fulldata$prop_is_branded == 1] <- "Yes"
fulldata$prop_is_branded[fulldata$prop_is_branded == 0] <- "No"
fulldata$prop_is_branded <- as.factor(fulldata$prop_is_branded)
fulldata$is_booking[fulldata$is_booking == 1] <- "Booked"
fulldata$is_booking[fulldata$is_booking == 0] <- "Did not book"
fulldata$is_booking <- as.factor(fulldata$is_booking) 
```

#### Relabel the categories of price, popularity, and distance
```{r}
# store the initial values
temp_price <- fulldata$hist_price_band
temp_pop <- fulldata$popularity_band
temp_dist <- fulldata$distance_band

# create two dictionaries to convert values
dict1 <- c(
  "VL" = "Very Low",
  "L" = "Low",
  "M" = "Medium", 
  "H" = "High", 
  "VH" = "Very High"
)
dict2 <- c(
  "VC" = "Very Close",
  "C" = "Close",
  "M" = "Medium",
  "F" = "Far",
  "VF" = "Very Far"
)

# convert values
fulldata$hist_price_band <- dict1[temp_price]
fulldata$popularity_band <- dict1[temp_pop]
fulldata$distance_band <- dict2[temp_dist]

# define the order of values
fulldata$hist_price_band <- factor(fulldata$hist_price_band, levels = c("Very Low", "Low", "Medium", "High", "Very High"))
fulldata$popularity_band <- factor(fulldata$popularity_band, levels = c("Very Low", "Low", "Medium", "High", "Very High"))
fulldata$distance_band <- factor(fulldata$distance_band, levels = c("Very Far", "Far", "Medium", "Close", "Very Close"))
```

# Create New Variables

#### Compute the number of days users plan to stay
```{r}
fulldata$stay_length <- as.numeric(difftime(fulldata$srch_co, fulldata$srch_ci, units = "days"))
```

#### Whether the user is traveling to a different country
```{r}
fulldata$abroad <- factor(NA, levels = c("Domestic", "International"))
fulldata$abroad[fulldata$user_location_country == fulldata$hotel_country] <- "Domestic"
fulldata$abroad[fulldata$user_location_country != fulldata$hotel_country] <- "International"
```

#### Whether the user is from the US
```{r}
fulldata$User_Location <- factor(NA, levels = c(
  "US", "Not US"
))

dict_location <- function(x) {
  if(x == "UNITED STATES OF AMERICA") {
    return("US")
  }
  else {
    return("Not US")
  }
}

fulldata$User_Location <- apply(matrix(fulldata$user_location_country), 1, dict_location)
```

#### Season
```{r}
fulldata$month <- month(as.Date(fulldata$date_time))

# a dictionary to convert values into seasons based on months
dict_season <- function(month){
  if (month >= 3 & month < 6) {
    return("Spring")
  }
  if (month >= 6 & month < 9) {
    return("Summer")
  }
  if (month >= 9 & month < 12) {
    return("Fall")
  }
  return("Winter")
}

fulldata$season <- apply(matrix(fulldata$month), 1, dict_season)
```

#### Popularity Score (Computed using the average of travel-related facet scores)
```{r}
fulldata$pop_score <- fulldata %>% 
  select(starts_with("popular_")) %>%
  apply(1, function(x) {as.numeric(sum(x) * -1 - 300)})
```
# Final Clean Up

```{r}
# remove irrelevant columns
fulldata <- fulldata[, substr(colnames(fulldata), 1, 8) != "popular_"]

# remove irrational & missing data
fulldata$stay_length[fulldata$stay_length < 0] <- NA
fulldata$orig_destination_distance[fulldata$orig_destination_distance < 0] <- NA
fulldata <- fulldata %>% drop_na(stay_length, orig_destination_distance, pop_score)
```

# Plots 
### (Note some plots are conditionally displayed based on the variable chosen)

#### How is distance related to the number of children and adults traveling?
```{r}
fit_adult <- lm(orig_destination_distance ~ srch_adults_cnt, data = fulldata)
fit_children <- lm(orig_destination_distance ~ srch_children_cnt, data = fulldata)

plot_adult <- plot_ly(fulldata, x = ~srch_adults_cnt) %>%
      add_markers(y = ~orig_destination_distance) %>%
      add_lines(x = ~srch_adults_cnt, y = fitted(fit_adult)) %>%
      layout(xaxis = list(title = "Number of Adults"), 
             yaxis = list(title = "User Distance from The Booked Hotel"),
             showlegend = F)
plot_children <- plot_ly(fulldata, x = ~srch_children_cnt) %>%
      add_markers(y = ~orig_destination_distance) %>%
      add_lines(x = ~srch_children_cnt, y = fitted(fit_children))  %>%
      layout(xaxis = list(title = "Number of Childen"),
             showlegend = F)
subplot(plot_adult, plot_children, nrows = 1, shareY = T, shareX = T) %>%
      layout(title = "<b> Distance <b>")
```

#### How is length of stay related to the number of children and adults traveling?
```{r}
fit_adult <- lm(stay_length ~ srch_adults_cnt, data = fulldata)
fit_children <- lm(stay_length ~ srch_children_cnt, data = fulldata)

plot_adult <- plot_ly(fulldata, x = ~srch_adults_cnt) %>%
  add_markers(y = ~stay_length) %>%
  add_lines(x = ~srch_adults_cnt, y = fitted(fit_adult)) %>%
  add_text(x = 10, y = 10, text = "100") %>%
  layout(xaxis = list(title = "Number of Adults"), 
         yaxis = list(title = "User Intended Length of Stay"), 
         showlegend = F)
plot_children <- plot_ly(fulldata, x = ~srch_children_cnt) %>%
  add_markers(y = ~stay_length) %>%
  add_lines(x = ~srch_children_cnt, y = fitted(fit_children))  %>%
  layout(xaxis = list(title = "Number of Childen"),
         showlegend = F)
subplot(plot_adult, plot_children, nrows = 1, shareY = T, shareX = T) %>%
  layout(title = "<b> Length <b>")
```

#### How do hotel preferences differ for domestic and international travelers? (Conditionally Displayed)
```{r, message = FALSE}
exp_var <- c("is branded hotel", "hotel star rating", "hotel price band", "hotel popularity band")[3] # choose explanatory variable

if (exp_var == "is branded hotel") {
  table <- fulldata %>%
    group_by(abroad, prop_is_branded) %>%
    summarize(count = n()) %>%
    mutate(prop = count / sum(count))
  
  plot_ly(table, x = ~abroad, y = ~prop, color = ~prop_is_branded) %>%
    add_bars() %>%
    layout(xaxis = list(title = "Hotel Location"), 
           yaxis = list(title = "Percentage of Users"),
           legend = list(title = list(text = "<b> Whether the Hotel Is Branded <b>")),
           title = list(text = "<b> Is Branded Hotel <b>", x = 0.3))
} else if (exp_var == "hotel star rating") {
  table <- fulldata %>%
    group_by(abroad, prop_starrating_cat) %>%
    summarize(count = n()) %>%
    mutate(prop = count / sum(count))
  
  plot_ly(table, x = ~abroad, y = ~prop, color = ~prop_starrating_cat) %>%
    add_bars() %>%
    layout(xaxis = list(title = "Hotel Location"), 
           yaxis = list(title = "Percentage of Users"),
           legend = list(title = list(text = "<b> Hotel Star Rating <b>")),
           title = list(text = "<b> Hotel Star Rating <b>", x = 0.3))
} else if (exp_var == "hotel price band") {
  table <- fulldata %>%
    group_by(abroad, hist_price_band) %>%
    summarize(count = n()) %>%
    mutate(prop = count / sum(count))
  
  plot_ly(table, x = ~abroad, y = ~prop, color = ~hist_price_band) %>%
    add_bars() %>%
    layout(xaxis = list(title = "Hotel Location"), 
           yaxis = list(title = "Percentage of Users"),
           legend = list(title = list(text = "<b> Hotel Price Ragne <b>")),
           title = list(text = "<b> Hotel Price Band <b>", x = 0.3))
}else if (exp_var == "hotel popularity band") {
  table <- fulldata %>%
    group_by(abroad, popularity_band) %>%
    summarize(count = n()) %>%
    mutate(prop = count / sum(count))
  
  plot_ly(table, x = ~abroad, y = ~prop, color = ~popularity_band) %>%
    add_bars() %>%
    layout(xaxis = list(title = "Hotel Location"), 
           yaxis = list(title = "Percentage of Users"),
           legend = list(title = list(text = "<b> Hotel Popularity <b>")),
           title = list(text = "<b> Hotel Popularity Band <b>", x = 0.3))
}
```

#### Popular Destinations for US and Non-US Users
```{r}
world <- map_data("world")

ggplot() +
  geom_polygon(data = world, aes(x = long, y = lat, group = group),
               fill = "white", color = "black") +
  geom_point(data = fulldata, aes(x = srch_destination_longitude,
                                  y = srch_destination_latitude,
                                  color = User_Location),
             alpha = 0.5) +
  scale_color_manual(values = c("red", "blue")) + 
  ggtitle("Popular Destinations of US and Non-US Users") +
  theme(axis.text = element_blank(),
        axis.title = element_blank(),
        axis.ticks = element_blank()) 


non_us <- fulldata %>% filter(User_Location == "US")

ggplot() +
  geom_polygon(data = world, aes(x = long, y = lat, group = group),
               fill = "white", color = "black") +
  geom_point(data = non_us, aes(x = srch_destination_longitude,
                                  y = srch_destination_latitude,
                                  color = User_Location),
             alpha = 0.5, color = "blue") + 
  ggtitle("Popular Destinations of US and Non-US Users") +
  theme(axis.text = element_blank(),
        axis.title = element_blank(),
        axis.ticks = element_blank()) 


us <- fulldata %>%
  filter(User_Location == "Not US")

ggplot() +
  geom_polygon(data = world, aes(x = long, y = lat, group = group),
               fill = "white", color = "black") +
  geom_point(data = us, aes(x = srch_destination_longitude,
                                  y = srch_destination_latitude,
                                  color = User_Location),
             alpha = 0.5, color = "red") + 
  ggtitle("Popular Destinations of US and Non-US Users") +
  theme(axis.text = element_blank(),
        axis.title = element_blank(),
        axis.ticks = element_blank()) 
```

#### How does percent booking change across seasons? 
```{r, message = FALSE, warning = FALSE}
season_tab <- fulldata %>%
      group_by(season, is_booking) %>%
      summarize(n = n()) %>%
      mutate(prop = n / sum(n))
    
    plot_ly(season_tab, x = ~season, y = ~prop, color = ~factor(is_booking)) %>%
      add_bars() %>%
      layout(xaxis = list(title = "Season"),
             yaxis = list(title = "Percent of Users"),
             title = "Booking by Season")
```

#### How does the number of bookings/clicks vary for different times in the day?
```{r, message = FALSE}
bookings_tab <- fulldata %>%
      group_by(hour, is_booking) %>%
      summarize(count = n()) %>%
      filter(is_booking == "Booked")
      
    plot_ly(bookings_tab, x = ~hour, y = ~count, type = "scatter", mode = "line") %>%
      layout(xaxis = list(title = "Hour of Day"),
             yaxis = list(title = "Number of Bookings"), 
             title = "Booking by Hour of Day")
```

#### How do clicks and bookings vary for different hotels? (Conditional Display)
```{r, message = FALSE, warnings = FALSE}
exp_var <- c("is branded hotel", "hotel star rating", "hotel price band", "hotel popularity band")[4] # choose explanatory variable

clicks <- fulldata %>% filter(is_booking == "Did not book")
response <- "Clicks"

if (exp_var == "is branded hotel") {
  explanatory <- "Brand"
  tab <- clicks %>%
    group_by(prop_is_branded, is_booking) %>%
    summarize(count = n())
  
  plot_ly(tab, x = ~prop_is_branded, y = ~count, color = ~prop_is_branded) %>%
    add_bars() %>%
    layout(xaxis = list(title = "Is Branded Hotel"),
           yaxis = list(title = paste("Number of ", response)),
           title = paste(response, "by", explanatory), 
           showlegend = F)
} else if (exp_var == "hotel star rating") {
  explanatory <- "Star Rating"
  tab <- clicks %>%
    group_by(prop_starrating_cat, is_booking) %>%
    summarize(count = n())
  
  plot_ly(tab, x = ~prop_starrating_cat, y = ~count, color = ~prop_starrating_cat) %>%
    add_bars() %>%
    layout(xaxis = list(title = "Hotel Star Rating"),
           yaxis = list(title = paste("Number of ", response)),
           title = paste(response, "by", explanatory),
           showlegend = F)
} else if (exp_var == "hotel price band") {
  explanatory <- "Price Band"
  tab <- clicks %>%
    group_by(hist_price_band, is_booking) %>%
    summarize(count = n())
  
  plot_ly(tab, x = ~hist_price_band, y = ~count, color = ~hist_price_band) %>%
    add_bars() %>%
    layout(xaxis = list(title = "Hotel Price Range"),
           yaxis = list(title = paste("Number of ", response)),
           title = paste(response, "by", explanatory),
           showlegend = F)
} else if (exp_var == "hotel popularity band") {
  explanatory <- "Popularity Band"
  tab <- clicks %>%
    group_by(popularity_band, is_booking) %>%
    summarize(count = n())
  
  plot_ly(tab, x = ~popularity_band, y = ~count, color = ~popularity_band) %>%
    add_bars() %>%
    layout(xaxis = list(title = "Hotel Popularity Band"),
           yaxis = list(title = paste("Number of ", response)),
           title = paste(response, "by", explanatory), 
           showlegend = F)
}

bookings <- fulldata %>% filter(is_booking == "Booked")
response <- "Bookings"
if (exp_var == "is branded hotel") {
  explanatory <- "Brand"
  tab <- bookings %>%
    group_by(prop_is_branded, is_booking) %>%
    summarize(count = n())
  
  plot_ly(tab, x = ~prop_is_branded, y = ~count, color = ~prop_is_branded) %>%
    add_bars() %>%
    layout(xaxis = list(title = "Is Branded Hotel"),
           yaxis = list(title = paste("Number of ", response)),
           title = paste(response, "by", explanatory), 
           showlegend = F)
} else if (exp_var == "hotel star rating") {
  explanatory <- "Star Rating"
  tab <- bookings %>%
    group_by(prop_starrating_cat, is_booking) %>%
    summarize(count = n())
  
  plot_ly(tab, x = ~prop_starrating_cat, y = ~count, color = ~prop_starrating_cat) %>%
    add_bars() %>%
    layout(xaxis = list(title = "Hotel Star Rating"),
           yaxis = list(title = paste("Number of ", response)),
           title = paste(response, "by", explanatory),
           showlegend = F)
} else if (exp_var == "hotel price band") {
  explanatory <- "Price Band"
  tab <- bookings %>%
    group_by(hist_price_band, is_booking) %>%
    summarize(count = n())
  
  plot_ly(tab, x = ~hist_price_band, y = ~count, color = ~hist_price_band) %>%
    add_bars() %>%
    layout(xaxis = list(title = "Hotel Price Range"),
           yaxis = list(title = paste("Number of ", response)),
           title = paste(response, "by", explanatory),
           showlegend = F)
} else if (exp_var == "hotel popularity band") {
  explanatory <- "Popularity Band"
  tab <- bookings %>%
    group_by(popularity_band, is_booking) %>%
    summarize(count = n())
  
  plot_ly(tab, x = ~popularity_band, y = ~count, color = ~popularity_band) %>%
    add_bars() %>%
    layout(xaxis = list(title = "Hotel Popularity Band"),
           yaxis = list(title = paste("Number of ", response)),
           title = paste(response, "by", explanatory), 
           showlegend = F)
}
```

#### How does popularity/price/distance affect the likeability of hotels? (Conditional Display)
```{r, message = FALSE, warnings = FALSE}
res_var <- c("length of stay", "hotel star rating", "travel-related facets score")[2] # choose response variable
exp_var <- c("popularity band", "price band", "distance band")[2] # choose explanatory variable

if (res_var == "length of stay") {
  res_name <- "Length of Stay"
  if (exp_var == "popularity band") {
    exp_name <- "Popularity Band"
    tab <- fulldata %>%
      group_by(popularity_band) %>%
      summarize(avg = mean(stay_length))
    
    plot_ly(tab, x = ~popularity_band, y = ~avg, color = ~popularity_band) %>%
      add_bars() %>%
      layout(xaxis = list(title = exp_name),
             yaxis = list(title = res_name), 
             title = paste(res_name, "by", exp_name),
             showlegend = F)
  }
  else if (exp_var == "price band") {
    exp_name <- "Price Range"
    tab <- fulldata %>%
      group_by(hist_price_band) %>%
      summarize(avg = mean(stay_length))
    
    plot_ly(tab, x = ~hist_price_band, y = ~avg, color = ~hist_price_band) %>%
      add_bars() %>%
      layout(xaxis = list(title = exp_name),
             yaxis = list(title = res_name), 
             title = paste(res_name, "by", exp_name),
             showlegend = F)
  }
  else if (exp_var == "distance band") {
    exp_name <- "Distance Band"
    tab <- fulldata %>%
      group_by(distance_band) %>%
      summarize(avg = mean(stay_length))
    
    plot_ly(tab, x = ~distance_band, y = ~avg, color = ~distance_band) %>%
      add_bars() %>%
      layout(xaxis = list(title = exp_name),
             yaxis = list(title = res_name), 
             title = paste(res_name, "by", exp_name),
             showlegend = F)
  }
} else if (res_var == "hotel star rating") {
  res_name <- "Hotel Star Rating"
  if (exp_var == "popularity band") {
    exp_name <- "Popularity Band"
    tab <- fulldata %>%
      group_by(popularity_band) %>%
      summarize(avg = mean(prop_starrating))
    
    plot_ly(tab, x = ~popularity_band, y = ~avg, color = ~popularity_band) %>%
      add_bars() %>%
      layout(xaxis = list(title = exp_name),
             yaxis = list(title = res_name), 
             title = paste(res_name, "by", exp_name),
             showlegend = F)
  }
  else if (exp_var == "price band") {
    exp_name <- "Price Range"
    tab <- fulldata %>%
      group_by(hist_price_band) %>%
      summarize(avg = mean(prop_starrating))
    
    plot_ly(tab, x = ~hist_price_band, y = ~avg, color = ~hist_price_band) %>%
      add_bars() %>%
      layout(xaxis = list(title = exp_name),
             yaxis = list(title = res_name), 
             title = paste(res_name, "by", exp_name),
             showlegend = F)
  }
  else if (exp_var == "distance band") {
    exp_name <- "Distance Band"
    tab <- fulldata %>%
      group_by(distance_band) %>%
      summarize(avg = mean(prop_starrating))
    
    plot_ly(tab, x = ~distance_band, y = ~avg, color = ~distance_band) %>%
      add_bars() %>%
      layout(xaxis = list(title = exp_name),
             yaxis = list(title = res_name), 
             title = paste(res_name, "by", exp_name),
             showlegend = F)
  }
} else if (res_var == "travel-related facets score") {
  res_name <- "Travel-Related Facets Score"
  if (exp_var == "popularity band") {
    exp_name <- "Popularity Band"
    tab <- fulldata %>%
      group_by(popularity_band) %>%
      summarize(avg = mean(pop_score))
    
    plot_ly(tab, x = ~popularity_band, y = ~avg, color = ~popularity_band) %>%
      add_bars() %>%
      layout(xaxis = list(title = exp_name),
             yaxis = list(title = res_name), 
             title = paste(res_name, "by", exp_name),
             showlegend = F)
  }
  else if (exp_var == "price band") {
    exp_name <- "Price Range"
    tab <- fulldata %>%
      group_by(hist_price_band) %>%
      summarize(avg = mean(pop_score))
    
    plot_ly(tab, x = ~hist_price_band, y = ~avg, color = ~hist_price_band) %>%
      add_bars() %>%
      layout(xaxis = list(title = exp_name),
             yaxis = list(title = res_name), 
             title = paste(res_name, "by", exp_name),
             showlegend = F)
  }
  else if (exp_var == "distance band") {
    exp_name <- "Distance Band"
    tab <- fulldata %>%
      group_by(distance_band) %>%
      summarize(avg = mean(pop_score))
    
    plot_ly(tab, x = ~distance_band, y = ~avg, color = ~distance_band) %>%
      add_bars() %>%
      layout(xaxis = list(title = exp_name),
             yaxis = list(title = res_name), 
             title = paste(res_name, "by", exp_name),
             showlegend = F)
  }
}
```
